Analyzing the Citation Rate of a Scientific Paper Using ML and Graph Algorithms

### 2024-11-10 paper/paper.tex
Adapt scripts to new file structure
add clustering results to paper

\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{degree_distribution.png}
\end{figure}

The first distribution is rather typical for a scale-free network and follows the power law. However, it is not exactly the case with the out-degree distribution. The possible explanation to this is the following: if we look at the graph as a temporal graph (we have dates of each paper's publishing) out-degree of each node is determined on creation and does not change over time. In-degree, on the other hand, has a behaviour that is much closer to how degrees act in a usual social graph, developing over time.

@@

\subsection{Time Properties}
\tab Since our graph can be interpreted as a temporal graph, we can study the way its properties change over time. For instance, we discovered that the number of new papers per unit of time decreases:
\begin{figure}[h]
\centering
\includegraphics[width=0.9\linewidth]{new_papers_over_time.png}
\end{figure}
\newpage
Another interesting thing that we found is how average in-degrees and out-degrees change. This is the plot of average degrees in 2 months time from 1991 to 2003:
\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{degrees_over_time.png}
\end{figure}
A downwards trend in the end for in-degree and upwards trend in the beginning are noticeable. Explanation to this phenomena is rather simple: older papers had more time to gain popularity and therefore have more incoming edges, whereas papers from 2003 do not have any papers in our dataset that cite them. The opposite happens with the out-degree. Earlier papers cited even earlier papers (from 1980s) that were not included in our graph and thus have little to no outgoing edges. It is also worth noticing that the growth of out-degree slows down closer to 2003. It means that almost enough papers are included in the graph at that time to publish a paper sourcing only these modern papers in our dataset.

\end{document}

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{degree_distribution.png}
\end{figure}
\newpage
The first distribution is rather typical for a scale-free network and follows the power law. However, it is not exactly the case with the out-degree distribution. The possible explanation to this is the following: if we look at the graph as a temporal graph (we have dates of each paper's publishing) out-degree of each node is determined on creation and does not change over time. In-degree, on the other hand, has a behaviour that is much closer to how degrees act in a usual social graph, developing over time.

@@

\subsection{Time Properties}
\tab Since our graph can be interpreted as a temporal graph, we can study the way its properties change over time. For instance, we discovered that the number of new papers per unit of time decreases:
\begin{figure}[h]
\centering
\includegraphics[width=0.9\linewidth]{new_papers_over_time.png}
\end{figure}
\newpage
Another interesting thing that we found is how average in-degrees and out-degrees change. This is the plot of average degrees in 2 months time from 1991 to 2003:
\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{degrees_over_time.png}
\end{figure}
A downwards trend in the end for in-degree and upwards trend in the beginning are noticeable. Explanation to this phenomena is rather simple: older papers had more time to gain popularity and therefore have more incoming edges, whereas papers from 2003 do not have any papers in our dataset that cite them. The opposite happens with the out-degree. Earlier papers cited even earlier papers (from 1980s) that were not included in our graph and thus have little to no outgoing edges. It is also worth noticing that the growth of out-degree slows down closer to 2003. It means that almost enough papers are included in the graph at that time to publish a paper sourcing only these modern papers in our dataset.
\subsection{Cluster properties}
We compared Louvain and label propagation community detection algorithms. All other methods were extremely slow. The best modularity was achieved by Louvain algorithm: 0.655 and 31 communities with fixed seed compared to 969 classes and 0.541 modularity using label propagation. We have divided the articles into the same number of clusters according to their abstracts using text embedding from `glove-wiki-gigaword-50` model (we averaged the embeddings of each word in the abstract) and KMeans algorithm. We also tried aglomerative clustering and gaussian mixture. Modularity for all of them was really poor, which is not suprising. That's the visualisation of it:
\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{Aglomerative.png}
\end{figure}
\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{gm.png}
\end{figure}
\end{document}
